---
title: "Homework Assignment #4"
subtitle: "Marine Aquaculture Suitability by EEZ Zone on West Coast of the U.S."
author: "Nicole Pepper"
date: last-modified
execute: 
  eval: true
  warning: false
  message: false
  results: "hide"
format:
  html:
    toc: true 
    code-fold: true
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---
#### Set up Workspace
```{r}
#| code-fold: true
#| code-summary: "Display Code"

# ---- Load libraries ----
# Load libraries
library(terra)
library(geodata)
library(tidyverse)
library(tmap)
library(kableExtra)
library(spData)
library(spDataLarge)
library(here)
library(viridisLite) 
library(terra)
library(sf)
library(purrr)
library(stars)
library(raster)
library(dplyr)
library(testthat)
```

#### Read in Project Data:
```{r}
#| code-fold: true
#| code-summary: "Display Code"

# ---- Read in Project Data ----

# Create data folder object
data_folder <- here::here("data")

# Read in sea surface temperature (sst) data & make mean
sst_list <- list.files(data_folder,
                       pattern = "average_annual_sst_\\d{4}\\.tif$",
                       full.names = TRUE)
sst_stack <- lapply(sst_list, rast) %>%
  rast()

# Read in depth data
depth <- rast(here::here("data/depth.tif"))

# Read in EEZ Zones
eez_zones <- st_read(here::here("data/wc_regions_clean.shp"))

```

#### Create fancy reproject function for list of all data
use `class()`to separate SpatRaster and sf

#### Prepare Data
```{r}

# ---- Prepare SST Data ----

# Calculate mean SST
mean_sst <- mean(sst_stack)

# Convert SST from Kelvin to Celsius
mean_sst <- mean_sst - 273.15


# ---- Check CRS of data ----

# Create a warning if the data do not match
if (crs(depth) != crs(mean_sst)) {
  warning("The datasets do not have the same CRS.")
} else {
  message("All datasets are in the same CRS.")
}

# ---- Resample raster data to match resolution and extent ----

# Transform CRS
eez_zones <- st_transform(eez_zones, crs = crs(mean_sst)) 

# Resample depth data to match the resolution of SST with nearest neighbor
depth <- resample(depth, mean_sst, method = "near")

# Crop the resampled depth data to the extent of SST data
depth <- crop(depth, mean_sst, snap = "near")

# ---- Reproject depth to match mean_sst data ----

# Reproject depth layer
depth <- project(depth, crs(mean_sst))

# Create a warning if the data do not match
if (crs(depth) != crs(mean_sst)) {
  warning("The datasets do not have the same CRS.")
} else {
  message("All datasets are in the same CRS.")
}

# Stack data
data_stack <- c(depth, mean_sst)
names(data_stack) <- c("depth", "mean_sst")

```


Oyster Suitability Conditions:

- sea surface temperature: 11-30°C

- depth: 0-70 meters below sea level

```{r}
# Does this include/exclude things right?
oy_sst <- matrix(c(-Inf, 11, 0,
                11, 30, 1,
                30, Inf, 0),
              ncol = 3, byrow = TRUE)

oy_depth <- matrix(c(-Inf, 0, 0,
                0, 70, 1,
                70, Inf, 0),
              ncol = 3, byrow = TRUE)

oy_sst_zone <- terra::classify(data_stack[["mean_sst"]], rcl = oy_sst)
oy_depth_zone <- terra::classify(data_stack[["depth"]], rcl = oy_depth)

oyster_suitable = oy_sst_zone * oy_depth_zone
```


```{r}
custom_colors <- c("white", "red")

# Plot the raster with custom colors
plot(oyster_suitable, col = custom_colors)
```

Function arguments: `suitability_fun`

- `minimum and maximum sea surface temperature`

- `minimum and maximum depth`

- `species name`
# ```{r}
# sst_suitability <- function(raster_layer, sst_min, sst_max, species) {
#   if (species == "oyster") {
#     oy_sst <- matrix(c(-Inf, sst_min, 0,
#                 sst_min, sst_max, 1,
#                 sst_max, Inf, 0),
#               ncol = 3, byrow = TRUE)
#     
#     oy_sst_suitablility <- terra::classify(raster_layer[["mean_sst"]], rcl = oy_sst)
#     return(oy_sst_suitablility)
#   } else {
#     return(NULL)
#   }
# }
# 
# ```

```{r}
oyster_sst_suitability_zone <- sst_suitability(data_stack, 11, 17, "oyster")
```

```{r}

# Plot the raster with custom colors
plot(oyster_sst_suitability_zone, col = custom_colors)
```

oy_depth <- matrix(c(-Inf, 0, 0,
                0, 70, 1,
                70, Inf, 0),
              ncol = 3, byrow = TRUE)
              
```{r}
habitat_suitability <- function(raster_layer, sst_min, sst_max, depth_min, depth_max, species) {
  if (species == "oyster") {
    oy_sst <- matrix(c(-Inf, sst_min, 0,
                sst_min, sst_max, 1,
                sst_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    oy_sst_suitability <- terra::classify(raster_layer[["mean_sst"]], rcl = oy_sst)

    oy_depth <- matrix(c(-Inf, depth_min, 0,
                depth_min, depth_max, 1,
                depth_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    oy_depth_suitability <- terra::classify(raster_layer[["depth"]], rcl = oy_depth)
    
    oyster_suitability <- oy_sst_suitability * oy_depth_suitability
    return(oyster_suitability)
  } 
  
else if (species == "giant kelp") {
  sal_sst <- matrix(c(-Inf, sst_min, 0,
                sst_min, sst_max, 1,
                sst_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    sal_sst_suitability <- terra::classify(raster_layer[["mean_sst"]], rcl = sal_sst)

    sal_depth <- matrix(c(-Inf, depth_min, 0,
                depth_min, depth_max, 1,
                depth_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    sal_depth_suitability <- terra::classify(raster_layer[["depth"]], rcl = sal_depth)
    
    sal_suitability <- sal_sst_suitability * sal_depth_suitability
    return(sal_suitability)
  
}
    
  else {
    return(NULL)
  }
}
```

oy_sst <- matrix(c(-Inf, 11, 0,
                11, 30, 1,
                30, Inf, 0),
              ncol = 3, byrow = TRUE)

oy_depth <- matrix(c(-Inf, 0, 0,
                0, 70, 1,
                70, Inf, 0),
              ncol = 3, byrow = TRUE)
### Take 2
# l apply

```{r}
habitat_suitability_map <- function(species, min_sst, max_sst, min_depth, max_depth) {
    # Assign suitable sst range
    sst <- matrix(c(-Inf, min_sst, 0,
                  min_sst, max_sst, 1,
                  max_sst, Inf, 0),
                ncol = 3, byrow = TRUE)
    sst_suitability <- terra::classify(data_stack[["mean_sst"]], rcl = sst)

    # Assign suitable depth range
  depth <- matrix(c(-Inf, min_depth, 0,
                        min_depth, max_depth, 1,
                       max_depth, Inf, 0),
                       ncol = 3, byrow = TRUE)
    
    depth_suitability <- terra::classify(data_stack[["depth"]], rcl = depth)

     # Reclassify combined suitability to ensure binary (0 or 1)
  binary_reclassification <- matrix(c(-Inf, 0, 0,  # Values < 0 become 0
                                       0, Inf, 1), # Values >= 0 become 1
                                     ncol = 3, byrow = TRUE)
  
    # Combine SST and depth suitability
    species_suitability <- sst_suitability * depth_suitability
    
    return(
  tm_shape(species_suitability) + 
  tm_raster(palette = c("#07222B", "#E0D71B"),
            labels = c("0 = Not Suitable", "1 = Suitable")) + 
    tm_layout(main.title = paste(species, "Habitat Suitability"),
              main.title.position = c("center", "top"),
              legend.text.color = "white",
              legend.position = c("bottom", "center"),
              bg.color = "#00BFB2")
)
}

```


```{r}
# ---- Create suitability function ----
habitat_suitability_map <- function(species, min_sst, max_sst, min_depth, max_depth) {

# ---- Assign suitable sst range ----
    sst <- matrix(c(-Inf, min_sst, 0,
                  min_sst, max_sst, 1,
                  max_sst, Inf, 0),
                ncol = 3, byrow = TRUE)
    sst_suitability <- terra::classify(data_stack[["mean_sst"]],
                                       rcl = sst,
                                       include.lowest = TRUE)
# ---- Assign suitable depth range ----
    # Assign suitable depth range
  depth <- matrix(c(-Inf, min_depth, 0,
                        min_depth, max_depth, 1,
                       max_depth, Inf, 0),
                       ncol = 3, byrow = TRUE)
    
    depth_suitability <- terra::classify(data_stack[["depth"]],
                                         rcl = depth,
                                         include.lowest = TRUE)

# ---- Combine suitability ----
    # Combine SST and depth suitability
    species_suitability <- sst_suitability * depth_suitability
    
   # Separate unsuitable area, reclassify 1 to NA
   species_unsuitable <- species_suitability
   species_unsuitable[species_unsuitable == 1] <- NA 
   
   # Separate unsuitable area, reclassify 0 to NA
   species_suitability[species_suitability == 0] <- NA

    

# ---- Calculate Suitable Area By EEZ Zones ----

# Create a raster where each cell is it's size (will be used to summarize area)
    #suitable_eez
cell_area <- cellSize(species_suitability, unit = 'km', transform = TRUE)

# Rasterize eez data by rgn field based on species_suitability resolution and extent
eez_raster <- rasterize(eez_zones, species_suitability, field= 'rgn')

# Mask the eez_raster to the species_suitability area
eez_suitability <-  mask(eez_raster, species_suitability)

# Calculate the area.
eez_suitability_area <- zonal(cell_area, eez_suitability, fun = 'sum')

eez_suitability_area_calc <- left_join(eez_zones, eez_suitability_area, by = 'rgn') %>% 
  mutate(eez_suitability_area = area,
         percentage = (eez_suitability_area / area_km2) * 100,
         .before = geometry)

eez_suitable_rating <- data.frame(region = eez_suitability_area_calc$rgn, percent_suitable = eez_suitability_area_calc$percentage)


# ---- Create suitability map ----
eez_suitability_area_calc$rgn_label <- gsub(" ", "\n", eez_suitability_area_calc$rgn)

return(

  tm_shape(species_unsuitable) + 
    tm_raster(palette = "#3D6399") +
  
  tm_shape(eez_suitability_area_calc) +
    tm_fill(col = "percentage",
            palette = c("#B2EFEF", "#6BCFC8", "#00BFB2", "#008F8A", "#004A45"),
            title = "Suitability Percentage",
            style = "cont",
            border.alpha = 0.8) +
# add labels
tm_shape(eez_suitability_area_calc) +  
  tm_text("rgn_label",
          size = 0.6,
          col = "white",
          shadow = TRUE,
          fontface = "bold") + 
    
     tm_shape(species_suitability) +
    tm_raster(palette = "#E0D71B",
              labels = "Suitable Area",
              title = '') + 
  
  tm_layout(main.title = paste(species, "Habitat Suitability"),
            main.title.position = c("center", "top"),
            legend.text.color = "white",
            legend.position = c(0.05,0.01),
            bg.color = '#07222B')

)
}

```

("#B2EFEF", "#6BCFC8", "#00BFB2", "#008F8A", "#004A45")

Kelp:
min sst: 10
max sst: 20
min depth: 0
max depth: 40


Oyster: 
min sst: 11
max sst: 30
min depth: 0
max depth: 70


```{r}
habitat_suitability_map("Oyster", 11,30,-70,0)

#plot(suitable_kelp_habitat, col = custom_colors)
#plot(eez_zones, add = TRUE, border = "black", lwd = 2)
```

```{r}
habitat_suitability_map("Kelp", 10,20,-40,0)
```

vOyster: 
min sst: 11
max sst: 30
min depth: 0
max depth: 70

```{r}
# ---- Assign suitable sst range ----
    sst <- matrix(c(-Inf, 11, 0,
                  11, 30, 1,
                  30, Inf, 0),
                ncol = 3, byrow = TRUE)
    sst_suitability <- terra::classify(data_stack[["mean_sst"]],
                                       rcl = sst,
                                       include.lowest = TRUE)
# ---- Assign suitable depth range ----
    # Assign suitable depth range
  depth <- matrix(c(-Inf, -70, 0,
                        -70, 0, 1,
                       0, Inf, 0),
                       ncol = 3, byrow = TRUE)
    
    depth_suitability <- terra::classify(data_stack[["depth"]],
                                         rcl = depth,
                                         include.lowest = TRUE)

# ---- Combine suitability ----
    # Combine SST and depth suitability
    species_suitability <- sst_suitability * depth_suitability
    

# ---- Calculate Suitable Area By EEZ Zones ----

# Calculate area
suitable_eez <- cellSize(species_suitability, unit = 'km', transform = TRUE)


# Rasterize eez data by rgn field
eez_raster<- rasterize(eez_zones, species_suitability, field= 'rgn')
eez_suitability <-  mask(eez_raster, species_suitability)

#calculate the area.
eez_suitability_area <- zonal(suitable_eez, eez_suitability, fun = 'sum')

eez_suitability_area_calc <- left_join(eez_zones, eez_suitability_area, by = 'rgn') %>% 
  mutate(eez_suitability_area = area,
         percentage = (eez_suitability_area / area_km2) * 100,
         .before = geometry)

eez_suitable_rating <- data.frame(region = eez_suitability_area_calc$rgn, percent_suitable = eez_suitability_area_calc$percentage)


# ---- Create suitability map ----
    return(
  tm_shape(species_suitability) + 
  tm_raster(palette = c("#07222B", "#E0D71B"),
            labels = c("0 = Not Suitable", "1 = Suitable"),
            title = '') + 
    
  tm_shape(eez_suitability_area_calc) +
    tm_fill(col = "percentage",
            palette = "-RdYlGn",  # Reverse the color palette for better visibility
            title = "Suitability Percentage",
            style = "cont",
            border.alpha = 0.5) +
    
    tm_layout(main.title = paste("Oyster", "Habitat Suitability"),
              main.title.position = c("center", "top"),
              legend.text.color = "white",
              legend.position = c("center", "bottom"),
              bg.color = "#00BFB2")
)
```

