---
title: "Homework Assignment #4"
subtitle: "Marine Aquaculture Suitability by EEZ Zone on West Coast of the U.S."
author: "Nicole Pepper"
date: last-modified
execute: 
  eval: true
  warning: false
  message: false
format:
  html:
    toc: true 
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| code-fold: true
#| code-summary: "Display Code"

# ---- Load libraries ----
# Load libraries
library(terra)
library(geodata)
library(tidyverse)
library(tmap)
library(kableExtra)
library(spData)
library(spDataLarge)
library(here)
library(viridisLite) 
library(terra)
library(sf)
library(purrr)
library(stars)
library(raster)
library(dplyr)

```

Read in Project Data:

```{r}
#| code-fold: true
#| code-summary: "Display Code"

# ---- Read in Project Data ----

# Create data folder object
data_folder <- here::here("data")

# Read in sea surface temperature (sst) data & make mean
sst_list <- list.files(data_folder, pattern = "average_annual_sst_\\d{4}\\.tif$", full.names = TRUE)
sst_stack <- lapply(sst_list, rast) %>% rast()

# Read in depth data
depth <- rast(here::here("data/depth.tif"))

# Read in EEZ Zones
eez_zones <- st_read(here::here("data/wc_regions_clean.shp"))

```

#### Create fancy reproject function for list of all data
use `class()`to separate SpatRaster and sf

#### Process Data
```{r}

# ---- Prepare SST Data ----

# Calculate mean SST
mean_sst <- mean(sst_stack)

# Convert SST from Kelvin to Celsius
mean_sst <- mean_sst - 273.15


# ---- Check CRS of data ----

# Create a warning if the data do not match
if (crs(depth) != crs(mean_sst)) {
  warning("The datasets do not have the same CRS.")
} else {
  message("All datasets are in the same CRS.")
}

# ---- Resample raster data to match resolution and extent ----

# Transform CRS
eez_zones <- st_transform(eez_zones, crs = crs(mean_sst)) 

# Resample depth data to match the resolution of SST with nearest neighbor
depth <- resample(depth, mean_sst, method = "near")

# Crop the resampled depth data to the extent of SST data
depth <- crop(depth, mean_sst, snap = "near")

# ---- Reproject depth to match mean_sst data ----

depth <- project(depth, crs(mean_sst))

# Create a warning if the data do not match
if (crs(depth) != crs(mean_sst)) {
  warning("The datasets do not have the same CRS.")
} else {
  message("All datasets are in the same CRS.")
}

# ---- Resample raster data to match resolution and extent ----
data_stack <- c(depth, mean_sst)
names(data_stack) <- c("depth", "mean_sst")

```


Oyster Suitability Conditions:

- sea surface temperature: 11-30°C

- depth: 0-70 meters below sea level

```{r}
# Does this include/exclude things right?
oy_sst <- matrix(c(-Inf, 11, 0,
                11, 30, 1,
                30, Inf, 0),
              ncol = 3, byrow = TRUE)

oy_depth <- matrix(c(-Inf, 0, 0,
                0, 70, 1,
                70, Inf, 0),
              ncol = 3, byrow = TRUE)

oy_sst_zone <- terra::classify(data_stack[["mean_sst"]], rcl = oy_sst)
oy_depth_zone <- terra::classify(data_stack[["depth"]], rcl = oy_depth)

oyster_suitable = oy_sst_zone * oy_depth_zone
```


```{r}
custom_colors <- c("white", "red")

# Plot the raster with custom colors
plot(oyster_suitable, col = custom_colors)
```

Function arguments: `suitability_fun`

- `minimum and maximum sea surface temperature`

- `minimum and maximum depth`

- `species name`
```{r}
sst_suitability <- function(raster_layer, sst_min, sst_max, species) {
  if (species == "oyster") {
    oy_sst <- matrix(c(-Inf, sst_min, 0,
                sst_min, sst_max, 1,
                sst_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    oy_sst_suitablility <- terra::classify(raster_layer[["mean_sst"]], rcl = oy_sst)
    return(oy_sst_suitablility)
  } else {
    return(NULL)
  }
}

```


```{r}
oyster_sst_suitability_zone <- oyster_suitability(data_stack, 11, 17, "oyster")
```

```{r}

# Plot the raster with custom colors
plot(oyster_sst_suitability_zone, col = custom_colors)
```

oy_depth <- matrix(c(-Inf, 0, 0,
                0, 70, 1,
                70, Inf, 0),
              ncol = 3, byrow = TRUE)
              
```{r}
habitat_suitability <- function(raster_layer, sst_min, sst_max, depth_min, depth_max, species) {
  if (species == "oyster") {
    oy_sst <- matrix(c(-Inf, sst_min, 0,
                sst_min, sst_max, 1,
                sst_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    oy_sst_suitability <- terra::classify(raster_layer[["mean_sst"]], rcl = oy_sst)

    oy_depth <- matrix(c(-Inf, depth_min, 0,
                depth_min, depth_max, 1,
                depth_max, Inf, 0),
              ncol = 3, byrow = TRUE)
    
    oy_depth_suitability <- terra::classify(raster_layer[["depth"]], rcl = oy_depth)
    
    oyster_suitability <- oy_sst_suitability * oy_depth_suitability
    return(oyster_suitability)
  } else {
    return(NULL)
  }
}
```

```{r}
suitable_oyster_habitat <- habitat_suitability(data_stack, 11, 17, 0, 70, "oyster")
plot(suitable_oyster_habitat, col = custom_colors)
```

